---
title: "R Notebook"
output: html_notebook
---

# Libraries

```{r, include = FALSE}
library(tidyverse)
library(microbiome)
library(phyloseq)
library(caret)
library(randomForest)
library(MButils)
```

# Load data

I am using the AHA dataset for writing my code because it is the simplest dataset I can work with.
Simplest = significant number of baseline samples without issue of repeated measures, balanced distribution of classes, likely difference to be detected present.


```{r}
p_ps = readRDS("/Users/aa370/Library/CloudStorage/Box-Box/project_davidlab/LAD_LAB_Personnel/Ammara_A/Projects/POMMS/R24_POMMS/data/processed/20221104_pomms_metadata_plant.rds")

p_ps

sample_data(p_ps)
```

# Preprocess

## Remove problematic samples

Taking out samples with less than or equal to 0 reads and keeping only entry timepoint reduces samples from 384 down to 240.

```{r}
p_ps = p_ps %>%
  subset_samples( reads > 0) %>%
  subset_samples(timepoint == "Entry")

p_ps
```

Class distribution in this dataset:
Case = 193
Control = 47

```{r}
sample_data(p_ps) %>%
  as.data.frame() %>%
  as_tibble() %>%
  group_by(group) %>%
  summarise(count = n())
```

## Data transform

```{r}
otu_clr = abundances(p_ps, "clr") %>% # rclr transform could not be performed, get Nans
  t()

p_ps = phyloseq(otu_table(otu_clr, taxa_are_rows = FALSE),
                   sample_data(sample_data(p_ps)),
                   tax_table(tax_table(p_ps)))

rm(otu_clr)
```

## Feature table

### Extract data

Order of asv colnames in OTU table and tax table is the same.

```{r}
features = p_ps@otu_table %>%
  as.data.frame() %>%
  as_tibble(rownames = NA) %>%
  rownames_to_column("sample")

feature_labels = c("sample")

 names = tax_table(p_ps) %>%
      data.frame() %>%
      lowest_level() %>%
   pull(name)

 feature_labels = append(feature_labels, names)

 feature_labels = make.unique(feature_labels) %>%
   make.names()

 colnames(features) = feature_labels

features
```
### Attach data labels

```{r}
sample_labels = p_ps@sam_data%>%
  as.data.frame() %>%
  as_tibble(rownames = NA) %>%
  rownames_to_column("sample") %>%
  select(sample, group)

features = sample_labels %>%
  left_join(features)
```

### Zero variance features

No features have zero variance
```{r}
nzv = features %>%
  select(!c(sample,group)) %>%
  nearZeroVar(saveMetrics = TRUE)
```

```{r}
rm(nzv)
```

### Correlated predictors

```{r}
feature_cor = features %>%
  select(!c(sample,group)) %>%
 cor() 

summary(feature_cor[upper.tri(feature_cor)])
```

```{r}
high_cor = findCorrelation(feature_cor, cutoff = .75)

filtered_features = features[,-high_cor]
```

```{r}
rm(high_cor)
rm(feature_cor)
```

### Linear dependencies

```{r}
feature_ld = filtered_features %>%
  select(!c(sample,group)) %>%
  findLinearCombos()

filtered_features = filtered_features[, - feature_ld$remove]
```

## Input table

Preprocessing reduces column number to 131 from 235

```{r}
input = filtered_features %>%
  mutate(group_shuffle = sample(group) %>%
           as.factor(),
         group = as.factor(group))
```

# Modeling

Load functions from classification_function_AA file

```{r}
label_list = c("group", "group_shuffle")
results = loop_label(iterations =20, input, label_list, percent = 0.7, downsampling = TRUE)

auc_df = results[[2]]

importance_df = results[[1]]
```

```{r}
rm(label_list)
```

# Plotting

## AUC plot and t-test
```{r}
auc_df %>%
  pivot_longer(cols = c("group", "group_shuffle"), names_to = "feature", values_to = "AUC") %>%
  ggplot()+
  geom_boxplot(aes(x=feature, y = AUC))+
  theme_classic()

t.test(auc_df$group, auc_df$group_shuffle)

```

## Importance plots

### look-up table
```{r}
#Making asv reference for names in model
lookup_asv = tax_table(p_ps) %>% 
     data.frame() %>%
      lowest_level() %>%
   rownames_to_column("asv") %>% 
   select(asv, name) %>%
   mutate(name = make.names(make.unique(name)))
```

### summarize importances
```{r}
# Making summary df of importances
importance_summary = importance_df %>%
  pivot_longer(cols = c("NA.":"Apium.graveolens"), names_to = "food", values_to = "importance") %>%
  group_by(variable, food) %>%
  summarise(mean_importance = mean(importance)) %>%
  arrange(variable, desc(mean_importance))
```
### top 10 ranked list
```{r}
# Getting the ranked list
ranking = importance_summary %>%
  filter(variable == "group") %>%
  pull(food)

top10 = ranking[1:10]
```

### Importance plot
```{r}
importance_df %>%
  pivot_longer(cols = c("NA.":"Apium.graveolens"), names_to = "food", values_to = "importance")  %>%
  filter(variable == "group" & food %in% top10) %>%
  ggplot()+
  geom_boxplot(aes(x= reorder(food,-importance), y = importance))+
  labs(x= "Food", y ="Importance") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust=1))
```

Some of the important NAs for predictions only map to bacterial species in BLAST.

# Important taxa directions

```{r}
sample_data(p_ps) %>%
  data.frame() %>%
  filter(timepoint == "Entry") %>%
  ggplot() +
  geom_boxplot(aes(x= group, y = bmi))+
  theme_classic()

plot_food_direction(top10, input, variable = "group")
```


```{r}
lookup_asv %>%
  filter(name == "NA.") %>%
  pull(asv)
```


```{r}
feature_cor %>%
  as.data.frame() %>%
  select(Helianthus.annuus) %>%
  arrange(desc(Helianthus.annuus))
```


